---
title: "Geography 176A"
author: "[Van Jackson](https://van-jackson.github.io)"
subtitle: 'Lab 05: Rasters & Remote Sensing'
output:
  html_document:
    theme: journal
---

```{r}
library(raster)
library(tidyverse)
library(getlandsat) 
library(sf) 
library(mapview)
library(osmdata)
library(factoextra)
library(leaflet)
```


#Question 1
```{r}
bb = read_csv('../data/uscities.csv') %>% 
  filter(city == "Palo") %>% 
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>% 
  st_transform(5070) %>% 
  st_buffer(5000) %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_as_sf()

```

#Question 2
```{r}
bbwgs = bb %>% st_transform(4326)
bb = st_bbox((bbwgs))

q2_image = getlandsat::lsat_scenes() %>% 
  filter(min_lat <= bb$ymin, min_lon <= bb$xmin) %>% 
  filter(max_lat >= bb$ymax, max_lon >= bb$xmax) %>% 
  filter(as.Date(acquisitionDate) == as.Date("2016-09-26"))

write.csv(q2_image, file = "../data/palo-flood-scene.csv")
```
### Step 2
```{r}
md = read_csv("../data/palo-flood-scene.csv")

files = lsat_scene_files(md$download_url) %>% 
  filter(grepl(paste0("B", 1:6, ".TIF$", collapse= "|"), file)) %>% 
  arrange(file) %>% 
  pull(file)

st= sapply(files,lsat_image)

s=stack(st) %>% setNames(c(paste0("band", 1:6)))


cropper= bbwgs %>%  st_transform(crs(s))

r=crop(s, cropper)

```



#Question 3
True 
```{r}
par(mfrow=c(1,2))
plotRGB(r, r=4 ,g=3 ,b=2, stretch= "lin")
plotRGB(r, r=4 ,g=3 ,b=2, stretch= "hist")
```

Infrared
```{r}
plotRGB(r, r=5 ,g=4 ,b=3, stretch= "lin")
plotRGB(r, r=5 ,g=4 ,b=3, stretch= "hist")
```

False Color 
```{r}
plotRGB(r, r=5 ,g=6 ,b=4, stretch= "lin")
plotRGB(r, r=5 ,g=6 ,b=4, stretch= "hist")
```

My Choice
```{r}
plotRGB(r, r=2 ,g=4 ,b=6, stretch= "lin")
plotRGB(r, r=2 ,g=4 ,b=6, stretch= "hist")
```
Stretching determines the representation of each value. Histogram stretching tends to be lighter and brighter, while linear stretching seems to be darker



# Question 4
5 Different Masks
```{r}
palette= colorRampPalette(c("blue","white","red"))

ndvi=(r$band5 - r$band4) / (r$band5 + r$band4)
ndwi=(r$band3 - r$band5) / (r$band3 + r$band5)
mndwi=(r$band3 - r$band6) / (r$band3 + r$band6)
wri=(r$band3 + r$band4) / (r$band5 + r$band6)
swi= 1 / sqrt(r$band2 - r$band6)
```

Plotting 
```{r}
plot(ndvi,col=palette(256), legend=FALSE)
plot(ndwi,col=palette(256), legend=FALSE)
plot(mndwi,col=palette(256), legend=FALSE)
plot(wri,col=palette(256), legend=FALSE)
plot(swi,col=palette(256), legend=FALSE)
```

Water Thresholds
```{r}
thresholding1= function(x){ifelse(x <= 0,1,NA)}
thresholding2= function(x){ifelse(x >= 0,1,NA)}
thresholding3= function(x){ifelse(x >= 0,1,NA)} 
thresholding4= function(x){ifelse(x >= 1,1,NA)}
thresholding5= function(x){ifelse(x <= 5,1,NA)}

flood1= calc(ndvi,thresholding1)
flood2= calc(ndwi,thresholding2)
flood3= calc(mndwi,thresholding3)
flood4= calc(wri,thresholding4)
flood5= calc(swi,thresholding5)

plot(flood1, col= "blue", legend=FALSE)
plot(flood2, col= "blue", legend=FALSE)
plot(flood3, col= "blue", legend=FALSE)
plot(flood4, col= "blue", legend=FALSE)
plot(flood5, col= "blue", legend=FALSE)

```

# Question 5
Kmeans
```{r}
set.seed(1)

r_extract = getValues(r)
r_nna = na.omit(r_extract) 


kmeans_r = kmeans(r_extract, 12)
fviz_cluster(kmeans_r, geom="point", data = r_extract)

thresholding_s= function(x){ifelse(x >= 0,1,0)}
flood_s= calc(ndwi,thresholding_s)
kmeans_raster = flood_s
values(kmeans_raster) = kmeans_r$cluster
plot(kmeans_raster, col = viridis::viridis(12))
```

Finding the correct KMeans cluster that represents our water pixels
```{r}
com_table = table(values(flood_s), values(kmeans_raster))

which.max(com_table[2,])

kmeans_raster[kmeans_raster != which.max(com_table[2,])] = 0
kmeans_raster[kmeans_raster != 0] =1
plot(kmeans_raster)
```

Changing remaining NAs to 0 for analysis
```{r}
thresholdings1= function(x){ifelse(x <= 0,1,0)}
thresholdings3= function(x){ifelse(x >= 0,1,0)} 
thresholdings4= function(x){ifelse(x >= 1,1,0)}
thresholdings5= function(x){ifelse(x <= 5,1,0)}

floods1= calc(ndvi,thresholdings1)
floods3= calc(mndwi,thresholdings3)
floods4= calc(wri,thresholdings4)
floods5= calc(swi,thresholdings5)

fs = stack(floods1, flood_s, floods3, floods4, floods5, kmeans_raster)
plot(fs)

sum_fs = fs %>% sum() 
plot(sum_fs, col = blues9)

(cellStats(fs, sum) * res(fs)^2) /1e6
```
# Extra Credit
How many layers classified the marked area in the video as a flood? 

```{r}
aoi = st_point(c(-91.78967, 42.06290)) %>% 
  st_sfc(crs = 4326) %>% 
  st_sf() %>% 
  st_transform(st_crs(fs))

raster::extract(sum_fs, aoi)
```
There are 3 layers that classified this area as a flood; it is assumed that some rasters missed the marking of the area as a flood
